---
title: "Graphical options with `plot_fullsibQTL`"
date: "`r Sys.Date()`"
author: "[Statistical Genetics Lab](http://statgen.esalq.usp.br) <br/> Department of Genetics <br/> Luiz de Queiroz College of Agriculture <br/> University of São Paulo"
output: 
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
vignette: >
  %\VignetteIndexEntry{Graphical options with plot_fullsibQTL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## From ggplot2 rmd vignette options
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "#>", 
                      fig.width = 7, 
                      fig.height = 5, 
                      fig.align = "center")
```

In this tutorial we show how to plot different graphics with the `plot_fullsibQTL` function. A complete description can be viewed through `?plot_fullsibQTL`. As an example we will use the example data from the package:

```{r}
  library(fullsibQTL)
  data("example_QTLfullsib")
  fullsib <- create_fullsib(example_QTLfullsib,
                             list(LG1_final, LG2_final, LG3_final, LG4_final),
                             step=1,map.function="kosambi",condIndex=3.5)

  ## IM
  im <- im_scan(fullsib)
  qtls.im <- r2_ls(fullsib, pheno.col=1, lg=c(1,1,2,2,2,3,3,4),
                 pos=c("loc13","M12","loc45","loc77","loc108","M33","loc65","M52"))
  ## CIM
  cofs.fs <- cof_selection(fullsib, pheno.col=1, k=log(300), n.cofactor = 10)
  cim<-cim_scan(cofs.fs)
  qtls.cim <- r2_ls(fullsib, pheno.col=1, lg=c(1,1,2,2,2,3,3,4),
                    pos=c("loc13","M12","loc45","loc77","loc108","M33","loc65","M52"))
  
  ## Getting im and cim peaks for alpha 0.05. 
  ## between () for created and print the object
  im.peak1 <- summary(im_perm,alpha=0.05)[1,1]
  im.peak2 <- summary(im_perm,alpha=0.05)[1,2]
  
  cim.peak1 <- summary(cim_perm, alpha=0.05)[1,1]
  cim.peak2 <- summary(cim_perm, alpha=0.05)[1,2]
```
## Customizing Graphics

The minimal plot has to have the fullsib object and at list one result of `fullsib_scan`. At the bottom of the graphic it is shown the genetic map.
```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=cim)
```

We can set the legend of the QTL mapping with the `qtlmapping` argument:
```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=cim,
                qtlmapping="CIM")
```

## Graphical comparision between different `(c)im_scan`
There is the option to insert multiple results from `(c)im_scan` with listing them with `list()`. Those multiple results can be different traits, a same trait in different environments, comparision of IM $\times$ CIM, and so on. Here we compare IM $\times$ CIM

```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                qtlmapping=c("CIM","IM"))
```

## Adding QTL location to the graphic
You can insert the QTLs position from the `r2_ls` function for each one of the `(c)im_scan`. In order to do it, you need to use the argument `r2ls.out`:

```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=cim,
                r2ls.out=qtls.cim,
                qtlmapping="CIM")
```

For multiple QTLs mappings:

```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"))
```

To set threshold lines, based on the `summary(cim_perm)` peaks 1 and 2 with $\alpha=0.05) results:

```{r}
summary(cim_perm)

plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906))
```

For printing purposes, another option is to change the graphic for a grayscale:

```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906),
                grayscale=TRUE)
```

There is the option to set which linkage groups you want to plot. To plot just LG 2:

```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906),
                lgs=2)
```

To plot LG 2 and 4:
```{r}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906),
                lgs=c(2,4))
```

## Interactive plot
There is also the option to plot a interactive graphic. Inside the function, we use the package `plotly` as the engine. With the option `interact=TRUE` it will be created an interactive `html` graphic in you working directory folder or you can set the folder name to create the `html`. The plot will be returned in your default browser. If you do not want to open the browser, set `browser=FALSE`

```{r, eval=FALSE}
## Setting the file name
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906), 
                interact=TRUE, 
                file="cim_im.html")

## Setting the folder name
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906), 
                interact=TRUE, 
                file="cim_im.html",
                folder="/home/your_user_name/Pictures")

## Setting to not open the browser
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906), 
                interact=TRUE, 
                file="cim_im.html",
                folder="/home/your_user_name/Pictures",
                browser=FALSE)
```

The above script creates the following graphic. It is a interactive `plotly`. *Click on legend entries to toggle traces, click-and-drag on the chart to zoom, double-click to autoscale, shift-and-drag to pan* ([plotly website](https://plot.ly/r/getting-started/)). Additionaly, it is included in the tooltip the LOD score, the QTL estimated effect, and marker names in the map.

<div>
<iframe src='http://statgen.esalq.usp.br/fullsibQTL_graphics/cim_im.html', height="600", width = "650"></iframe>
</div>

The user can also set the height and width of the interact plot within the arguments `height` and `width`, they are the same as the arguments of `plotly` function. The default of `height` is 500 pixels and for `width` is 250 $\times$ the number of LGs to be plot. All the arguments can be combined in the interactive plot. 

## Changing colors

There is the option to choose the colors for each LG, the argument `lg.colors` will pass its values to the `ggplot2::scale_colour_manual` function. You can check more colors options at this [cheatsheet](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf) or at this color [guide](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf).

Here we plot an interactive plot in order to compare the CIM and IM mapping setting two threshold values, `darkgreen` color for CIM and `gree` for IM, and choosing to plot just LG 2 and 3.

```{r, eval=FALSE}
plot_fullsibQTL(fullsib=fullsib,
                fullsib.scan=list(cim,im),
                r2ls.out=list(qtls.cim,qtls.im),
                qtlmapping=c("CIM","IM"),
                thr = c(4.023724, 2.900906), 
                lgs = c(2,4),
                interact=TRUE, 
                lg.colors=c("darkgreen","green"),
                height=600,
                width=800,
                file="cim_im_greens.html",
                folder="/home/your_user_name/Pictures")
```

<div>
<iframe src='http://statgen.esalq.usp.br/fullsibQTL_graphics/cim_im_greens.html', height="700", width = "650"> </iframe>
</div>

# Appendix C - Graphical options with plot
<a name="appendixc"> </a>
In this section we show how to plot different graphics with the `plot` function. A complete description can be viewed through `?plot.fullsib_scan`. As an example we will use the example data from the package:

```{r}
  library(fullsibQTL)
  data("example_QTLfullsib")
  fullsib <- create_fullsib(example_QTLfullsib,
                             list(LG1_final, LG2_final, LG3_final, LG4_final),
                             step=1,map.function="kosambi",condIndex=3.5)

  ## IM
  im <- im_scan(fullsib)
  qtls.im <- r2_ls(fullsib, pheno.col=1, lg=c(1,1,2,2,2,3,3,4),
                 pos=c("loc13","M12","loc45","loc77","loc108","M33","loc65","M52"))
  ## CIM
  cofs.fs <- cof_selection(fullsib, pheno.col=1, k=log(300), n.cofactor = 10)
  cim<-cim_scan(cofs.fs)
  qtls.cim <- r2_ls(fullsib, pheno.col=1, lg=c(1,1,2,2,2,3,3,4),
                    pos=c("loc13","M12","loc45","loc77","loc108","M33","loc65","M52"))

```

The basic graphic


You can check more colors options at this [cheatsheet](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf) or at this color [guide](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) 

## Customizing Graphics
```{r}
plot(im)

## argument lty
## Changes the line type
## There is a total of 6 different line types as follow:
plot(im)
abline(h=c(1,2,3,4,5,6),lty=c(1,2,3,4,5,6))

## argument lwd
## Changes the line width
## It is changed in a relative proportion to the default. 2 is twice the default width, 3 is triple of the default, and so on.
plot(im)
abline(h=c(.5,1,2,3,4,5,6),lwd=c(.5,1,2,3,4,5,6))

## You can set any of them with the plot with:
plot(im, lty=3, lwd=2)

## argument lg
## Plotting just LG1
plot(im, lty=3, lwd=2, lg=1)

## Plotting LG1 and LG2
plot(im, lty=3, lwd=2, lg=c(1,2))

## argument label.lg
## Labels the LGs
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"))

## argument cex.axis  
## Changes the magnification of axis annotation relative to cex
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1)
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1.5)

## argument incl.mkr
## Includes the marker position
## There are 3 options: `points`, `name`, and, `none`. `none` is default.
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1.2, incl.mkr="points")

## argument cex.incl
## Changes the magnification of incl.mrk annotation relative to cex.
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1.2, incl.mkr="name", cex.incl = .75)

## argument gap
## Enlarges the gap between chromosomes (in CM)
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1.2, incl.mkr="name", cex.incl = .75, gap = 50)

## argument col
## Changes the line colors
plot(im, lty=3, lwd=2, lg=c(1,2), label.lg=c("Chr 1","Chr 2"), cex.axis=1.2, incl.mkr="name", cex.incl = .75, gap = 50, col="red")
```

You can check more colors options at this [cheatsheet](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf) or at this color [guide](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) 

You can also customize the plot with the default R plot options as `main`, `ylab`, `xlab`, `yaxt`, `ylim`, etc. 

```{r}
plot(im, lty=1, lwd=2, lg=c(1,2), cex.axis=1.2, incl.mkr="name", cex.incl = .9, gap = 50, col="red", alternate.lgid = TRUE,
     ## Configuring with R plot options:
     main = "IM - Trait - Crop", 
     xlab = "Linkage Group", 
     ylab = "LOD score", 
     yaxt = "n",
     ylim = c(0,10))
axis(side=2,at=seq(0,10,1),lwd=1)
abline(h=im.peak1, lty=2)
abline(h=im.peak2, lty=3)
legend("topleft", lty=c(1,2,3,0), cex=0.75, col=c("red","black","black"), legend=c("IM","1st Peak Threshold","2nd Peak Threshold"))
```

## Graphical comparision between `(c)im_scan`
The IM and CIM mapping results can be compared in a graphical way with the following codes:
```{r}
plot(im,cim, lty = c(2,1), lwd = 2, incl.mkr = "name", cex.incl = 0.7, gap=50,
     cex.axis = 0.8, col = c("blue","red"),
     main = "IM - Trait - Crop", 
     xlab = "Linkage Group", 
     ylab = "LOD score", 
     yaxt = "n",
     ylim = c(0,15))
axis(side=2,at=seq(0,15,1),lwd=1)
abline(h = im.peak1, lty = 2, lwd = 0.75)
abline(h = cim.peak1, lty = 2, lwd = 0.75)
legend("topleft", lty=c(2,1,2,2), cex=0.5, col=c("red","blue","black","black"), legend=c("IM","CIM","1st Threshold","2nd Threshold"))


## Adding QTLs
```
## Adding QTL location to the graphic
Our QTLs are the following:
```{r}
print(cim,lg=1,pos=c("loc13","M12"))
print(cim,lg=2,pos=c("loc45","loc77","loc108"))
print(cim,lg=3,pos=c("M33","loc65"))
print(cim,lg=4,pos="M52")
```
To add them to the graphic we can use the function `points` as follows. Knowing that the `gap` space is `50`, we can add this to the total size of each LG and get the position. The points will be localized at:
```{r}
points.loc <- c(13,110,45,77,108,20,65,60)
points.loc <- points.loc + c(0,0,
                             140+50,140+50,140+50,
                             140+50+140+50,140+50+140+50,
                             140+50+140+50+140+50)
```
Therefore:
```{r}
plot(im,cim, lty = c(2,1), lwd = 2, incl.mkr = "name", cex.incl = 0.5, gap=50,
     cex.axis = 0.8, col = c("blue","red"),
     main = "IM - Trait - Crop", 
     xlab = "Linkage Group", 
     ylab = "LOD score", 
     yaxt = "n",
     ylim = c(-2,15))
axis(side=2,at=seq(0,15,1),lwd=1)
abline(h = im.peak1, lty = 2, lwd = 0.75)
abline(h = cim.peak1, lty = 2, lwd = 0.75)
points(y=rep(0,8),x=points.loc, pch=17, col="dark green")
legend("topleft", lty=c(2,1,2,2,0), cex=0.75, pch = c(NA,NA,NA,NA,17),col = c("red", "blue", "black", "black", "dark green"), legend = c("IM","CIM","1st Threshold","2nd Threshold","QTL"),bty='n',y.intersp=.75)
```

# Session Info

```{r}
sessionInfo()
```